---
title: "02_behav_&_parameter_analyses"
author: "Milena Musial"
format:
  html:
    embed-resources: true
    editor: visual
    toc: true
    toc-depth: 4
    number-sections: true
    number-depth: 4
---

# Preps

```{r}
rm(list=ls())

# install.packages('tidyverse', 'stringr')
libs<-c("tidyverse", "stringr", "plyr", "dplyr", "lme4", "robustlmm", "lmerTest",
        "sjPlot", "remef")
sapply(libs, require, character.only=TRUE)

cbPalette <- c( "#0072B2", "#D55E00", "#009E73", "#CC79A7",
                "#F0E442", "#56B4E9", "#999999", "#E69F00")


data_path<-'/fast/work/groups/ag_schlagenhauf/B01_FP1_WP2/ILT_DATA'

load(file.path(data_path,"Behav/behav_final_n71.RData"))
load(file.path(data_path,"RedCap/redcap_n71.RData"))
load(file.path(data_path,"Behav/behav_final_redcap_n71.RData"))

data_path_comp<-'C:/Users/musialm/OneDrive - Charité - Universitätsmedizin Berlin/PhD/04_B01/ILT/WP2_ILT_CODE/Stan Modeling/Output'

load(file.path(data_path_comp,"Params.RData"))

# create outcome as predictor for stay switch analyses
behav_final$outcome_pred <- factor(behav_final$outcome)
behav_final_redcap$outcome_pred <- factor(behav_final_redcap$outcome)

# code aud criteria and audit criteria as factor
#behav_final_redcap$aud_sum <- factor(behav_final_redcap$aud_sum)
#behav_final_redcap$audit_sum <- factor(behav_final_redcap$audit_sum)
#redcap$aud_sum <- factor(redcap$aud_sum)
#redcap$audit_sum <- factor(redcap$audit_sum)

# set contrasts
contrasts(behav_final$reinforcer_type) <- contr.treatment(2)
contrasts(behav_final_redcap$reinforcer_type) <- contr.treatment(2)

redcap$aud_group <- factor(redcap$aud_group, levels = c("HC", "AUD"))
contrasts(redcap$aud_group) <- contr.treatment(2)
contrasts(behav_final_redcap$aud_group) <- contr.treatment(2)

contrasts(behav_final$outcome_pred) <- contr.treatment(2)
contrasts(behav_final_redcap$outcome_pred) <- contr.treatment(2)

# se function
se <- function(x) sd(x, na.rm=TRUE)/sqrt(length(x)-sum(is.nan(x)))
```

# Mandatory exclusion criteria

Exclude subjects where 2) signal loss or artefacts in MRI images were visually detected OR 3) participants exhibited translation \> 4 mm or rotation \> 4° across each block

```{r}
ID_excl <- factor()
```

# Optional exclusion criteria

## More than 10% NaN per block

```{r}
df_NaN <- behav_final %>%
  group_by(ID,reinforcer_type) %>%
  summarise_at(vars(rt), funs(s_NaN = sum(is.na(.)), p_NaN = sum(is.na(.))/50)) %>%
  mutate(excl_rt = p_NaN>0.1)

# add IDs to optionally excluded IDs
ID_excl <- append(ID_excl,c(df_NaN$ID[df_NaN$p_NaN>0.1]))
```

Subject 12654 has 12 % missing responses and will optionally be excluded (both blocks).

## Same key in \>95% of trials

```{r}
df_key <- behav_final %>%
  group_by(ID,reinforcer_type) %>%
  summarise_at(vars(side), funs(p_left = sum(side==1, na.rm=T)/50)) %>%
  mutate(excl_side = p_left>0.95)

# add IDs to optionally excluded IDs
ID_excl <- append(ID_excl,c(df_key$ID[df_key$p_left>0.95]))
```

None of the subjects pressed the same key in \>95% of trials.

## 5% lowest taste ratings across blocks and participants

### Prepare dfs

```{r}
# get df with taste and craving ratings per subject and block
behav_rating <- behav_final %>%
  group_by(ID,reinforcer_type) %>%
  select(ID,reinforcer_type,ID_block,taste1, taste2, crave1, crave2) %>%
  distinct() %>%
  mutate(taste_diff=taste2-taste1,
         crave_diff=crave2-crave1)
behav_rating <- merge(behav_rating,redcap,by="ID",all.x = T) # merge with demographic data for linear mixed-effects regression


behav_rating_alc <- behav_rating[behav_rating$reinforcer_type=='A',]
behav_rating_jui <- behav_rating[behav_rating$reinforcer_type=='J',]
```

```{r}
behav_rating_lowest <- behav_rating[behav_rating$taste1 < quantile(behav_rating$taste1, 0.05),]

# add IDs to optionally excluded IDs
ID_excl <- append(ID_excl,c(behav_rating_lowest$ID))
```

8 subjects have taste1 scores \< the 5% percentile.

# Pleasantness and craving ratings

## Descriptive stats

```{r}
d_rating <- behav_rating %>%
  group_by(reinforcer_type) %>%
  dplyr::summarise(across(c('taste1','taste2','crave1','crave2','taste_diff','crave_diff'), .fns=
                   list('_mean' = mean,
                        '_sd' = sd,
                        '_min' = min,
                        '_max' = max),
                   na.rm=T))

d_rating_nan <- behav_rating %>%
  group_by(reinforcer_type) %>%
  dplyr::summarise(across(c('taste1','taste2','crave1','crave2','taste_diff','crave_diff'), 
                          .fns=list('_na' = ~sum(is.na(.)))))
d_rating_nan

# alc
stats <- d_rating[1,-1] %>% gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "__") %>%
  spread(stat, val) %>%
  mutate(across(c('taste1','taste2','crave1','crave2'), round, 2))
stats

# jui
stats <- d_rating[2,-1] %>% gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "__") %>%
  spread(stat, val) %>%
  mutate(across(c('taste1','taste2','crave1','crave2'), round, 2))
stats
```

## Linear regression

### Difference post-pre taste ratings

#### AUD group

```{r}
m_taste <- lm(taste_diff ~ 1 + reinforcer_type*aud_group, data=behav_rating)

tab_model(m_taste,
          dv.labels=c("m_taste"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

#### AUD criteria

```{r}
m_taste_audsum <- lm(taste_diff ~ 1 + reinforcer_type*aud_sum, data=behav_rating)

tab_model(m_taste_audsum,
          dv.labels=c("m_taste"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

#### AUDIT criteria

```{r}
m_taste_auditsum <- lm(taste_diff ~ 1 + reinforcer_type*audit_sum, data=behav_rating)

tab_model(m_taste_auditsum,
          dv.labels=c("m_taste"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

### Difference post-pre craving ratings

#### AUD group

```{r}
m_crave <- lm(crave_diff ~ 1 + reinforcer_type*aud_group, data=behav_rating)

tab_model(m_crave,
          dv.labels=c("m_crave"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

#### AUD criteria

```{r}
m_crave_audsum <- lm(crave_diff ~ 1 + reinforcer_type*aud_sum, data=behav_rating)

tab_model(m_crave_audsum,
          dv.labels=c("m_crave"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

#### AUDIT criteria

```{r}
m_crave_auditsum <- lm(crave_diff ~ 1 + reinforcer_type*audit_sum, data=behav_rating)

tab_model(m_crave_auditsum,
          dv.labels=c("m_crave"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

## Plotting

### Taste difference

#### Per group and reinforcer type

```{r}
fig1 <- ggplot(behav_rating, aes(reinforcer_type, y = taste_diff, fill=reinforcer_type)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  facet_wrap(~aud_group, scales="fixed") +
  scale_x_discrete("Reinforcer type", labels=c('alcohol', 'juice')) +
  scale_y_continuous("Taste rating post-pre") +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig1
```

#### Per group

```{r}
fig1 <- ggplot(behav_rating, aes(aud_group, y = taste_diff, fill=aud_group)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_x_discrete("AUD group", labels=c('HC', 'AUD')) +
  scale_y_continuous("Taste rating post-pre") +
  scale_fill_manual("AUD group", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig1
```

### Craving difference

#### Per group and reinforcer type

```{r}
fig1 <- ggplot(behav_rating, aes(reinforcer_type, y = crave_diff, fill=reinforcer_type)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  facet_wrap(~aud_group, scales="fixed") +
  scale_x_discrete("Reinforcer type", labels=c('alcohol', 'juice')) +
  scale_y_continuous("Craving rating post-pre") +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig1
```

#### Per group

```{r}
fig1 <- ggplot(behav_rating, aes(aud_group, y = crave_diff, fill=aud_group)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_x_discrete("AUD group", labels=c('HC', 'AUD')) +
  scale_y_continuous("Craving rating post-pre") +
  scale_fill_manual("AUD group", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig1
```

# Correct choices

## prepare dfs

```{r}
# get p correct per subject and block for linear mixed-effects regression
df_c <- behav_final %>%
  group_by(ID,reinforcer_type) %>%
  summarise_at(vars(correct), list(p_correct = mean), na.rm=T)
df_c <- merge(df_c,redcap,by="ID") # merge with demographic data for linear mixed-effects regression

# get p correct per block and trial for plotting
df_c_block_trial <- behav_final %>%
  group_by(reinforcer_type,trial_block) %>%
  summarise_at(vars(correct), list(p_correct = mean), na.rm=T)

# get p correct per block and aud group and trial for plotting
df_c_block_trial_group <- behav_final_redcap %>%
  group_by(aud_group,reinforcer_type,trial_block) %>%
  summarise_at(vars(correct), list(p_correct = mean), na.rm=T)
```

## Descriptive stats

```{r}
# per group and reinforcer type
df_c_stats <- behav_final_redcap %>%
  group_by(reinforcer_type,aud_group) %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_mean' = mean,
                        '_sd' = sd,
                        '_min' = min,
                        '_max' = max),
                   na.rm=T))
df_c_stats

df_c_stats_nan <- behav_final_redcap %>%
  group_by(reinforcer_type,aud_group) %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_na' = ~sum(is.na(.)))))
df_c_stats_nan

# per group
df_c_stats <- behav_final_redcap %>%
  group_by(aud_group) %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_mean' = mean,
                        '_sd' = sd,
                        '_min' = min,
                        '_max' = max),
                   na.rm=T))

df_c_stats

df_c_stats_nan <- behav_final_redcap %>%
  group_by(aud_group) %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_na' = ~sum(is.na(.)))))

df_c_stats_nan

# per reinforcer type
df_c_stats <- behav_final_redcap %>%
  group_by(reinforcer_type) %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_mean' = mean,
                        '_sd' = sd,
                        '_min' = min,
                        '_max' = max),
                   na.rm=T))

df_c_stats

df_c_stats_nan <- behav_final_redcap %>%
  group_by(reinforcer_type) %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_na' = ~sum(is.na(.)))))

df_c_stats_nan

# overall
df_c_stats <- behav_final_redcap %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_mean' = mean,
                        '_sd' = sd,
                        '_min' = min,
                        '_max' = max),
                   na.rm=T))

df_c_stats

df_c_stats_nan <- behav_final_redcap %>%
  dplyr::summarise(across(c('correct'), .fns=
                   list('_na' = ~sum(is.na(.)))))

df_c_stats_nan


table(behav_final_redcap$correct)

```

## Linear mixed-effects regression

### AUD group

```{r}
m_correct_1 <- lmer(p_correct ~ 1 + reinforcer_type*aud_group + (1 | ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)
#m_correct_2 <- lmer(p_correct ~ 1 + reinforcer_type*aud_group + (1 + reinforcer_type || ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)
#m_correct_3 <- lmer(p_correct ~ 1 + reinforcer_type*aud_group + (1 + reinforcer_type | ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)

#anova(m_correct_1,m_correct_2,m_correct_3)

tab_model(m_correct_1,
          dv.labels=c("m_correct"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

We cannot include reinforcer type as random effect as the model would be unidentifiable.

We find - n.s. effects.

### AUD criteria

```{r}
m_correct_1_audsum <- lmer(p_correct ~ 1 + reinforcer_type*aud_sum + (1 | ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)
#m_correct_2_audsum <- lmer(p_correct ~ 1 + reinforcer_type*aud_sum + (1 + reinforcer_type || ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)
#m_correct_3_audsum <- lmer(p_correct ~ 1 + reinforcer_type*aud_sum + (1 + reinforcer_type | ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)

#anova(m_correct_1_audsum,m_correct_2_audsum,m_correct_3_audsum)

tab_model(m_correct_1_audsum,
          dv.labels=c("m_correct"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

We cannot include reinforcer type as random effect as the model would be unidentifiable.

We find - n.s. effects.

### AUDIT criteria

```{r}
m_correct_1_auditsum <- lmer(p_correct ~ 1 + reinforcer_type*audit_sum + (1 | ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)
#m_correct_2_auditsum <- lmer(p_correct ~ 1 + reinforcer_type*audit_sum + (1 + reinforcer_type || ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)
#m_correct_3_auditsum <- lmer(p_correct ~ 1 + reinforcer_type*audit_sum + (1 + reinforcer_type | ID), data=df_c, control=lmerControl(calc.derivs=FALSE), REML=F)

#anova(m_correct_1_auditsum,m_correct_2_auditsum,m_correct_3_auditsum)

tab_model(m_correct_1_auditsum,
          dv.labels=c("m_correct"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

We cannot include reinforcer type as random effect as the model would be unidentifiable.

We find - n.s. effects.

## Exploratory: Logistic mixed-effects regression

### AUD group

```{r}
m_correct_log_1 <- glmer(correct ~ 1 + reinforcer_type*aud_group + (1 | ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))
m_correct_log_2 <- glmer(correct ~ 1 + reinforcer_type*aud_group + (1 + reinforcer_type || ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))
m_correct_log_3 <- glmer(correct ~ 1 + reinforcer_type*aud_group + (1 + reinforcer_type | ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))

anova(m_correct_log_1,m_correct_log_2)
anova(m_correct_log_2,m_correct_log_3)

tab_model(m_correct_log_2,
          dv.labels=c("m_correct_log"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

We find - a sig. effect of reinforcer type.

### AUD criteria

```{r}
m_correct_log_1_audsum <- glmer(correct ~ 1 + reinforcer_type*aud_sum + (1 | ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))
m_correct_log_2_audsum <- glmer(correct ~ 1 + reinforcer_type*aud_sum + (1 + reinforcer_type || ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))
m_correct_log_3_audsum <- glmer(correct ~ 1 + reinforcer_type*aud_sum + (1 + reinforcer_type | ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))

anova(m_correct_log_1_audsum,m_correct_log_2_audsum)
anova(m_correct_log_2_audsum,m_correct_log_3_audsum)

tab_model(m_correct_log_2_audsum,
          dv.labels=c("m_correct_log"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

We find - n.s. effects.

### AUDIT criteria

```{r}
m_correct_log_1_auditsum <- glmer(correct ~ 1 + reinforcer_type*audit_sum + (1 | ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))
m_correct_log_2_auditsum <- glmer(correct ~ 1 + reinforcer_type*audit_sum + (1 + reinforcer_type || ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))
m_correct_log_3_auditsum <- glmer(correct ~ 1 + reinforcer_type*audit_sum + (1 + reinforcer_type | ID), data=behav_final_redcap, family = binomial, control=glmerControl(calc.derivs=FALSE))

anova(m_correct_log_1_auditsum,m_correct_log_2_auditsum)
anova(m_correct_log_2_auditsum,m_correct_log_3_auditsum)

tab_model(m_correct_log_2_auditsum,
          dv.labels=c("m_correct_log"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))
```

## Plotting

### Per group and reinforcer type

```{r}
fig1 <- ggplot(df_c, aes(reinforcer_type, y = p_correct, fill=reinforcer_type)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  facet_wrap(~aud_group, scales="fixed") +
  scale_x_discrete("Reinforcer type", labels=c('alcohol', 'juice')) +
  scale_y_continuous("% correct") +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig1
```

More correct choices for juice compared to alcohol reinforcer across diagnostic groups. Descriptively, the difference between alcohol and juice reinforcer is lower in the AUD compared to the HC group.

### Correct choices over course of task

#### Per block

```{r}
fig2 <- ggplot(df_c_block_trial, aes(x=trial_block, y=p_correct, group=reinforcer_type)) +
  geom_line(aes(color=reinforcer_type)) +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(1,2)]) +
  theme_bw(base_size=12)

fig2

```

-   P correct per block and trial across subjects.

-   Drops most likely due to probabilistic events - same across all subjects, different across versions, but partially overlapping (e.g. 7,8 in version A, 7,8,9 in version B).

-   p_correct quickly rises above 0.5 --\> subjects chose rewarded card more often --\> rewards aren't generally aversive

-   p_correct is generally higher for juice vs. alcohol reward

-   no string ceiling effect, but p_correct rises quickly

#### Per block and aud group

```{r}
fig3 <- ggplot(df_c_block_trial_group, aes(x=trial_block, y=p_correct, group=reinforcer_type)) +
  geom_line(aes(color=reinforcer_type)) +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(1,2)]) +
  theme_bw(base_size=12) +
  facet_wrap(~aud_group, scales="free_y")

fig3

```

# Win-Stay behavior: Repeating the same action that was rewarded in previous trial

## Prepare dfs

```{r}
# get p correct per subject and block for linear mixed-effects regression
df_win_stay <- behav_final %>%
  group_by(ID,reinforcer_type,na.rm=T) %>%
  summarise_at(vars(stay,win_stay), list(p = mean), na.rm=T) %>%
  na.omit()
df_win_stay <- merge(df_win_stay,redcap,by="ID") # merge with demographic data for linear mixed-effects 

# get p correct per block and trial for plotting
df_win_stay_trial <- behav_final %>%
  group_by(reinforcer_type,trial_block) %>%
  summarise_at(vars(win_stay), list(p = mean), na.rm=T)

# get p correct per block and aud group and trial for plotting
df_win_stay_trial_group <- behav_final_redcap %>%
  group_by(aud_group,reinforcer_type,trial_block) %>%
  summarise_at(vars(win_stay), list(p = mean), na.rm=T)
```

## Linear mixed-effects regression win_stay

### AUD group

```{r}
m_win_stay_1=lmer(win_stay_p~reinforcer_type*aud_group + (1 | ID),
                data=df_win_stay,
                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_2=lmer(win_stay_p~reinforcer_type*aud_group + (1 + reinforcer_type || ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_3=lmer(win_stay_p~reinforcer_type*aud_group + (1 + reinforcer_type | ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#anova(m_win_stay_1,m_win_stay_2,m_win_stay_3)

tab_model(m_win_stay_1,
          dv.labels=c("m_stay"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

### AUD criteria

```{r}
m_win_stay_1_audsum=lmer(win_stay_p~reinforcer_type*aud_sum + (1 | ID),
                data=df_win_stay,
                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_2_audsum=lmer(win_stay_p~reinforcer_type*aud_sum + (1 + reinforcer_type || ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_3_audsum=lmer(win_stay_p~reinforcer_type*aud_sum + (1 + reinforcer_type | ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#anova(m_win_stay_1_audsum,m_win_stay_2_audsum,m_win_stay_3_audsum)

tab_model(m_win_stay_1_audsum,
          dv.labels=c("m_stay"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

### AUDIT criteria

```{r}
m_win_stay_1_auditsum=lmer(win_stay_p~reinforcer_type*audit_sum + (1 | ID),
                data=df_win_stay,
                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_2_auditsum=lmer(win_stay_p~reinforcer_type*audit_sum + (1 + reinforcer_type || ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_3_auditsum=lmer(win_stay_p~reinforcer_type*audit_sum + (1 + reinforcer_type | ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#anova(m_win_stay_1_auditsum,m_win_stay_2_auditsum,m_win_stay_3_auditsum)

tab_model(m_win_stay_1_auditsum,
          dv.labels=c("m_stay"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

## Logistic mixed-effects regression

### AUD group

```{r}
m_win_stay_log_1=lmer(win_stay~reinforcer_type*aud_group + (1 | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_win_stay_log_2=lmer(win_stay~reinforcer_type*aud_group + (1 + reinforcer_type || ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_win_stay_log_3=lmer(win_stay~reinforcer_type*aud_group + (1 + reinforcer_type | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))

anova(m_win_stay_log_1,m_win_stay_log_2)
anova(m_win_stay_log_2,m_win_stay_log_3)
anova(m_win_stay_log_1,m_win_stay_log_3)

tab_model(m_win_stay_log_3,
          dv.labels=c("m_stay"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

Model 2 fails to converge. Model 3 is better than model 1.

We find a sig. effect of reinforcer type.

### AUD criteria

```{r}
m_win_stay_log_1_audsum=lmer(win_stay~reinforcer_type*aud_sum + (1 | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_win_stay_log_2_audsum=lmer(win_stay~reinforcer_type*aud_sum + (1 + reinforcer_type || ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_win_stay_log_3_audsum=lmer(win_stay~reinforcer_type*aud_sum + (1 + reinforcer_type | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))

anova(m_win_stay_log_1_audsum,m_win_stay_log_2_audsum)
anova(m_win_stay_log_2_audsum,m_win_stay_log_3_audsum)
anova(m_win_stay_log_1_audsum,m_win_stay_log_3_audsum)

tab_model(m_win_stay_log_3_audsum,
          dv.labels=c("m_stay"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

### AUDIT criteria

```{r}
m_win_stay_log_1_auditsum=lmer(win_stay~reinforcer_type*audit_sum + (1 | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_win_stay_log_2_auditsum=lmer(win_stay~reinforcer_type*audit_sum + (1 + reinforcer_type || ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_win_stay_log_3_auditsum=lmer(win_stay~reinforcer_type*audit_sum + (1 + reinforcer_type | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))

anova(m_win_stay_log_1_auditsum,m_win_stay_log_2_auditsum)
anova(m_win_stay_log_2_auditsum,m_win_stay_log_3_auditsum)
anova(m_win_stay_log_1_auditsum,m_win_stay_log_3_auditsum)

tab_model(m_win_stay_log_2_auditsum,
          dv.labels=c("m_stay"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

## Plotting

### Violin plot per aud group and reinforcer type

```{r}
fig4 <- ggplot(df_win_stay, aes(aud_group, y = win_stay_p, fill=aud_group)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  facet_wrap(~ reinforcer_type, scales="fixed") +
  scale_x_discrete("AUD group", labels=c('HC', 'AUD')) +
  scale_y_continuous("prob(repeat action rewarded in prev trial)") +
  scale_fill_manual("AUD_group", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig4

```

### Violin plot per AUD group

```{r}
fig5 <- ggplot(df_win_stay, aes(aud_group, y = win_stay_p, fill=aud_group)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_x_discrete("AUD group", labels=c('HC', 'AUD')) +
  scale_y_continuous("prob(repeat action rewarded in prev trial)") +
  scale_fill_manual("AUD_group", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig5
```

### Violin plot per reinforcer type

```{r}
fig6 <- ggplot(df_win_stay, aes(reinforcer_type, y = win_stay_p, fill=reinforcer_type)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_x_discrete("reinforcer type", labels=c('alcohol', 'juice')) +
  scale_y_continuous("prob(repeat action rewarded in prev trial)") +
  scale_fill_manual("reinforcer_type", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig6
```

### Win-stay behavior over course of task

#### Per block

```{r}
fig2 <- ggplot(df_win_stay_trial, aes(x=trial_block, y=p, group=reinforcer_type)) +
  geom_line(aes(color=reinforcer_type)) +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(1,2)]) +
  theme_bw(base_size=12)

fig2

```

#### Per block and aud group

```{r}
fig3 <- ggplot(df_win_stay_trial_group, aes(x=trial_block, y=p, group=reinforcer_type)) +
  geom_line(aes(color=reinforcer_type)) +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(1,2)]) +
  theme_bw(base_size=12) +
  facet_wrap(~aud_group, scales="fixed")

fig3

```

# No-Win-Switch behavior: Switching the action if it was not rewarded in previous trial

## Prepare dfs

```{r}
# get p correct per subject and block for linear mixed-effects regression
df_nowin_switch <- behav_final %>%
  group_by(ID,reinforcer_type,na.rm=T) %>%
  summarise_at(vars(stay,nowin_switch), list(p = mean), na.rm=T) %>%
  na.omit()
df_nowin_switch <- merge(df_nowin_switch,redcap,by="ID") # merge with demographic data for linear mixed-effects 

# get p correct per block and trial for plotting
df_nowin_switch_trial <- behav_final %>%
  group_by(reinforcer_type,trial_block) %>%
  summarise_at(vars(nowin_switch), list(p = mean), na.rm=T)

# get p correct per block and aud group and trial for plotting
df_nowin_switch_trial_group <- behav_final_redcap %>%
  group_by(aud_group,reinforcer_type,trial_block) %>%
  summarise_at(vars(nowin_switch), list(p = mean), na.rm=T)
```

## Linear mixed-effects regression

### AUD group

```{r}
m_nowin_switch_1=lmer(nowin_switch_p~reinforcer_type*aud_group + (1 | ID),
                data=df_nowin_switch,
                control=lmerControl(calc.derivs=FALSE))
#m_nowin_switch_2=lmer(nowin_switch_p~reinforcer_type*aud_group + (1 + reinforcer_type || ID),
#                data=df_nowin_switch,
#                control=lmerControl(calc.derivs=FALSE))
#m_nowin_switch_3=lmer(nowin_switch_p~reinforcer_type*aud_group + (1 + reinforcer_type | ID),
#                data=df_nowin_switch,
#                control=lmerControl(calc.derivs=FALSE))
#anova(m_nowin_switch_1,m_nowin_switch_2,m_nowin_switch_3)

tab_model(m_nowin_switch_1,
          dv.labels=c("m_switch"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

### AUD criteria

```{r}
m_nowin_switch_1_audsum=lmer(nowin_switch_p~reinforcer_type*aud_sum + (1 | ID),
                data=df_nowin_switch,
                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_2_audsum=lmer(win_stay_p~reinforcer_type*aud_sum + (1 + reinforcer_type || ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#m_win_stay_3_audsum=lmer(win_stay_p~reinforcer_type*aud_sum + (1 + reinforcer_type | ID),
#                data=df_win_stay,
#                control=lmerControl(calc.derivs=FALSE))
#anova(m_win_stay_1_audsum,m_win_stay_2_audsum,m_win_stay_3_audsum)

tab_model(m_nowin_switch_1_audsum,
          dv.labels=c("m_switch"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

### AUDIT criteria

```{r}
m_nowin_switch_1_auditsum=lmer(nowin_switch_p~reinforcer_type*audit_sum + (1 | ID),
                data=df_nowin_switch,
                control=lmerControl(calc.derivs=FALSE))
#m_nowin_switch_2_auditsum=lmer(nowin_switch_p~reinforcer_type*audit_sum + (1 + reinforcer_type || ID),
#                data=df_nowin_switch,
#                control=lmerControl(calc.derivs=FALSE))
#m_nowin_switch_3_auditsum=lmer(nowin_switch_p~reinforcer_type*audit_sum + (1 + reinforcer_type | ID),
#                data=df_nowin_switch,
#                control=lmerControl(calc.derivs=FALSE))
#anova(m_nowin_switch_1_auditsum,m_nowin_switch_2_auditsum,m_nowin_switch_3_auditsum)

tab_model(m_nowin_switch_1_auditsum,
          dv.labels=c("m_switch"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

## Logistic mixed-effects regression

### AUD group

```{r}
m_nowin_switch_log_1=lmer(nowin_switch~reinforcer_type*aud_group + (1 | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_nowin_switch_log_2=lmer(nowin_switch~reinforcer_type*aud_group + (1 + reinforcer_type || ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_nowin_switch_log_3=lmer(nowin_switch~reinforcer_type*aud_group + (1 + reinforcer_type | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))

anova(m_nowin_switch_log_1,m_nowin_switch_log_2)
anova(m_nowin_switch_log_2,m_nowin_switch_log_3)
anova(m_nowin_switch_log_1,m_nowin_switch_log_3)

tab_model(m_nowin_switch_log_3,
          dv.labels=c("m_switch"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

Model 2 fails to converge. Model 3 is better than model 1.

We find a sig. effect of reinforcer type.

### AUD criteria

```{r}
m_nowin_switch_log_1_audsum=lmer(nowin_switch~reinforcer_type*aud_sum + (1 | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_nowin_switch_log_2_audsum=lmer(nowin_switch~reinforcer_type*aud_sum + (1 + reinforcer_type || ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_nowin_switch_log_3_audsum=lmer(nowin_switch~reinforcer_type*aud_sum + (1 + reinforcer_type | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))

anova(m_nowin_switch_log_1_audsum,m_nowin_switch_log_2_audsum)
anova(m_nowin_switch_log_2_audsum,m_nowin_switch_log_3_audsum)
anova(m_nowin_switch_log_1_audsum,m_nowin_switch_log_3_audsum)

tab_model(m_nowin_switch_log_3_audsum,
          dv.labels=c("m_switch"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

### AUDIT criteria

```{r}
m_nowin_switch_log_1_auditsum=lmer(nowin_switch~reinforcer_type*audit_sum + (1 | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_nowin_switch_log_2_auditsum=lmer(nowin_switch~reinforcer_type*audit_sum + (1 + reinforcer_type || ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))
m_nowin_switch_log_3_auditsum=lmer(nowin_switch~reinforcer_type*audit_sum + (1 + reinforcer_type | ID),
                data=behav_final_redcap,
                control=lmerControl(calc.derivs=FALSE))

anova(m_nowin_switch_log_1_auditsum,m_nowin_switch_log_2_auditsum)
anova(m_nowin_switch_log_2_auditsum,m_nowin_switch_log_3_auditsum)
anova(m_nowin_switch_log_1_auditsum,m_nowin_switch_log_3_auditsum)

tab_model(m_nowin_switch_log_2_auditsum,
          dv.labels=c("m_switch"), digits=3, digits.re=2, 
          show.se=TRUE, show.stat=TRUE, show.df=TRUE, show.ci=FALSE,CSS = css_theme("cells"))

```

## Plotting

### Violin plot per aud group and reinforcer type

```{r}
fig4 <- ggplot(df_nowin_switch, aes(aud_group, y = nowin_switch_p, fill=aud_group)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  facet_wrap(~ reinforcer_type, scales="fixed") +
  scale_x_discrete("AUD group", labels=c('HC', 'AUD')) +
  scale_y_continuous("prob(switch action if not rewarded in prev trial)") +
  scale_fill_manual("AUD_group", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig4

```

### Violin plot per AUD group

```{r}
fig5 <- ggplot(df_nowin_switch, aes(aud_group, y = nowin_switch_p, fill=aud_group)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_x_discrete("AUD group", labels=c('HC', 'AUD')) +
  scale_y_continuous("prob(switch action if not rewarded in prev trial)") +
  scale_fill_manual("AUD_group", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig5
```

### Violin plot per reinforcer type

```{r}
fig6 <- ggplot(df_nowin_switch, aes(reinforcer_type, y = nowin_switch_p, fill=reinforcer_type)) +
  geom_violin() +
  geom_boxplot(width=0.1) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_x_discrete("reinforcer type", labels=c('alcohol', 'juice')) +
  scale_y_continuous("prob(switch action if not rewarded in prev trial)") +
  scale_fill_manual("reinforcer_type", values = cbPalette[c(6,8)]) +
  theme_bw(base_size=12) +
  theme(legend.position = "none")

fig6
```

### Win-stay behavior over course of task

#### Per block

```{r}
fig2 <- ggplot(df_nowin_switch_trial, aes(x=trial_block, y=p, group=reinforcer_type)) +
  geom_line(aes(color=reinforcer_type)) +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(1,2)]) +
  theme_bw(base_size=12)

fig2

```

#### Per block and aud group

```{r}
fig3 <- ggplot(df_nowin_switch_trial_group, aes(x=trial_block, y=p, group=reinforcer_type)) +
  geom_line(aes(color=reinforcer_type)) +
  scale_fill_manual("Reinforcer type", values = cbPalette[c(1,2)]) +
  theme_bw(base_size=12) +
  facet_wrap(~aud_group, scales="fixed")

fig3

```
