---
title: "Acces Stan fitting results & prepare parametric modulators for fMRI analysis"
author: "Milena Musial"
output: html_document
date: '2023-09-24'
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
    fig-width: 8
    fig-height: 6
    fig-format: svg
    editor_options: 
    chunk_output_type: console
---

# Setup

```{r}
rm(list=ls())
libs<-c("rstan", "gdata", "bayesplot", "stringr", "dplyr", "ggplot2")
sapply(libs, require, character.only=TRUE)

datapath <- '/fast/work/groups/ag_schlagenhauf/B01_FP1_WP2/ILT_Stan_Modeling'
out_path <- '/fast/work/groups/ag_schlagenhauf/B01_FP1_WP2/ILT_Stan_Modeling/Output'
behavpath <- '/fast/work/groups/ag_schlagenhauf/B01_FP1_WP2/ILT_DATA'

fit_file <- 'fit_n_142_2023-09-25_bandit2arm_delta_main_estimation0_delta0.99_stepsize0.1.rds'

fit <- readRDS(file.path(out_path, fit_file)) # Stan model output
load(file.path(behavpath,"Behav/behav_final_redcap_n71.RData")) # Behav and redcap data

color_scheme_set("mix-blue-pink")
```

Load data from prior predictive checks (estimation==0) or from model fitting (estimation==1)?
```{r}
estimation <- 0
```

# Output summary (contents of fit object)

-   mu_pr: raw, untransformed group-level means of learning rate (alpha) and inverse temperature (tau)

-   sigma: raw, untransformed group-level variance of learning rate (alpha) and inverse temperature (tau)

-   A_pr: raw, untransformed subject-level learning rates (alpha)

-   tau_pr: raw, untransformed subject-level inverse temperature (tau)

-   **A**: transformed subject-level learning rates (alpha), range 0-1

-   **tau**: transformed subject-level inverse temperature (tau), range 0-5

-   **mu_A**: group-level mean of transformed subject-level learning rates (alpha), range 0-1

-   **mu_tau**: group-level mean of transformed subject-level inverse temperature (tau), range 0-5

-   log_lik: subject.level log likelihood for model comparisons

-   y_pred: choice data generated from estimated parameters

-   PE_pred: PEs per subject and trial

-   ev_pred: Q-values for both choices per subject and trial

-   ev_chosen_pred: Q-values for chosen card per subject and trial

# Convergence / sampler diagnostics

## Summary statistics for parameters

### Across chains

```{r}
fit_sum <- summary(fit)

fit_sum

# overall summary
#print(fit_sum$summary)

# max Rhat for entire output including generated quantities
max(fit_sum$summary[,10])
max(fit_sum$summary[,10], na.rm=T)

# min effective sample size for entire output including generated quantities
min(fit_sum$summary[,9])
min(fit_sum$summary[,9], na.rm=T)
```

-   All `Rhat` values are \< 1.01, meaning that chains have converged! \`

-   All `n_eff` (measure of the effective sample size) are \> 1/100th (360) or 1/1000th (36) of number of iterations (3600). Anything over an `n_eff` of 100 is usually "fine"

### For each chain separately

No Rhat or n_eff given here. This is just done for reasons of completeness.

```{r}
print(fit_sum$c_summary) 
```

## Traceplots of model parameters

```{r}
traceplot(fit, pars="mu_A", inc_warmup=T)
traceplot(fit, pars="sigma[1]", inc_warmup=T)
traceplot(fit, pars="mu_pr[1]", inc_warmup=T) # untransformed mean alpha
traceplot(fit, pars="A_pr[100]", inc_warmup=T) # untransformed subject-level alpha


traceplot(fit, pars="mu_tau",inc_warmup=T)
traceplot(fit, pars="sigma[2]", inc_warmup=T)
traceplot(fit, pars="mu_pr[2]", inc_warmup=T) # untransformed mean tau
traceplot(fit, pars="tau_pr[100]", inc_warmup=T) # untransformed subject-level tau
```

Chains mix nicely.

## Acceptance rate and tree depth per chain

```{r}
sampler_params<- get_sampler_params(fit, inc_warmup=FALSE)

mean_accept_stat_by_chain<-sapply(sampler_params, function(x) mean(x[, "accept_stat__"]))
print(mean_accept_stat_by_chain)

max_treedepth_by_chain <- sapply(sampler_params, function(x) max(x[, "treedepth__"]))
print(max_treedepth_by_chain)
```

# Inspect posteriors

## Posterior densities & histograms

```{r}
stan_hist(fit, pars=c("mu_A",'mu_pr[1]',"mu_tau",'mu_pr[2]'))
```

## Mean parameter estimates and credible intervals

```{r}
plot(fit, pars=c("mu_A", "mu_pr[1]", "mu_tau", "mu_pr[2]"), show_density = FALSE, ci_level = 0.5, outer_level = 0.95, fill_color = "salmon")
```

## Probability that each parameter is over / under a certain value of interest

```{r}
posterior <- extract(fit)

sum(posterior$A>0.5)/length(posterior$A)
sum(posterior$tau>0.1)/length(posterior$tau)
```

# Prior / Posterior predictive checks

## Load real subject data (input for Stan)

```{r}
load(file.path(datapath,"Input/stan_data.RData"))

choice <- stan_data[["choice"]]

# only select trials in which all subjects made a choice
choice <- c(choice[,1:47])
```

-   142 blocks \* 50 trials matrix including -999 values

## Extract simulated values from generated quantities block

```{r}
y_pred_cat_log_rng <- as.matrix(fit, pars = "y_pred_cat_log_rng")

# create logical vector coding if a column should be included in y_pred
x <- logical(stan_data$T)
a <- c(1:(stan_data$N*47))
x[a] <- TRUE

# exclude 48:50th trial per subject
y_pred_cat_log_rng <- y_pred_cat_log_rng[,x]
```

-   3600 (iterations) \* 7100 (142 blocks \* 50 trials) matrix

## Compare estimates of summary statistics between choice and predicted choice

```{r}
ppc_stat(y = choice, yrep = y_pred_cat_log_rng, stat = function(y) mean(y, na.rm=T))
```

## Bar plot: count of choices and estimated choices

```{r}
ppc_bars(y = choice, yrep = y_pred_cat_log_rng)
```

# Extract parameter estimates

## Alpha per subject

```{r}
# mean across iterations of all chains
A_summary <- as.matrix(summary(fit, pars = "A")$summary[, c("mean")]) 
```

## Inverse temperature per subject

```{r}
# mean across iterations of all chains
tau_summary <- as.matrix(summary(fit, pars = "tau")$summary[, c("mean")]) 
```

## Save parameter estimates together with ID, reinforcer type, aud group information

```{r}

params <- behav_final_redcap %>%
  group_by(ID,aud_group,reinforcer_type,block) %>%
  summarise_at(vars(correct), list(p_correct = mean), na.rm=T)

params$alpha <- A_summary
params$invtemp <- tau_summary

params

if (estimation==1) {
  write.table(params, file=file.path(out_path, 'Params_main.txt'), append = FALSE, sep = " ", dec = ".", row.names = F, col.names = c("ID","aud_group","reinforcer_type","block","p_corr","alpha","invtemp"))
}
```

# Trial-by-trial quantities (can be used as fMRI parametric modulators)

## RPEs

### Extract as parametric modulators

```{r}
# mean across iterations of all chains
PE_summary <- as.matrix(summary(fit, pars = "PE_pred")$summary[, c("mean")])
PE_summary <- t(matrix(PE_summary, nrow = 50)) # trial x subject matrix

mods <- cbind(matrix(params$ID),PE_summary)
mods
```

### Plot over course of trials

```{r}
# mean across iterations of all chains
PE_summary <- as.data.frame(summary(fit, pars = "PE_pred")$summary[, c("mean")])
PE_summary$ID_block <- behav_final_redcap$ID_block
PE_summary$trial_block <- behav_final_redcap$trial_block
PE_summary <- PE_summary %>%
  rename(PE_pred = `summary(fit, pars = "PE_pred")$summary[, c("mean")]`) %>%
  mutate_at(c('PE_pred'), ~na_if(., -999))

# across subjects
PE_summary_per_trial <- PE_summary %>%
  group_by(trial_block) %>%
  summarise_at(vars(PE_pred), list(PE_mean = mean), na.rm=T)

fig_PE_per_trial <- ggplot(PE_summary_per_trial, aes(x=trial_block, y=PE_mean)) +
  geom_point() +
  geom_line() +
  theme_bw(base_size=12)

fig_PE_per_trial

# for single subjects
fig_PE_per_trial <- ggplot(PE_summary[PE_summary$ID_block==c('11475002','11475001'),], aes(x=trial_block, y=PE_pred, group=ID_block, color=ID_block)) +
  geom_point() +
  geom_line() +
  theme_bw(base_size=12)

fig_PE_per_trial
```

## Chosen Q

### Extract as parametric modulators

```{r}
# mean across iterations of all chains
Q_summary <- as.matrix(summary(fit, pars = "ev_chosen_pred")$summary[, c("mean")])
Q_summary <- t(matrix(Q_summary, nrow = 50)) # trial x subject matrix
Q_summary

mods <- cbind(mods,Q_summary)

if (estimation==1) {
  write.table(mods, file=file.path(out_path, 'fmri_param_modulators_main.txt'), append = FALSE, sep = " ", dec = ".", row.names = F, col.names = F)
}
```

### Plot over course of trials

Currently trying to plot Q values of better and worse card separately over course of trials.

```{r}
# mean across iterations of all chains
Q_summary <- as.data.frame(summary(fit, pars = "ev_pred")$summary[, c("mean")])
Q_summary$ID_block <- behav_final_redcap$ID_block
Q_summary$trial_block <- behav_final_redcap$trial_block
Q_summary <- Q_summary %>%
  rename(Q_pred = `summary(fit, pars = "ev_pred")$summary[, c("mean")]`) %>%
  mutate_at(c('Q_pred'), ~na_if(., -999))
Q_summary$row_odd <- seq_len(nrow(Q_summary)) %% 2

# incorrect card
Q_summary_1 <- Q_summary[Q_summary$row_odd==1,]

Q_summary_per_trial_1 <- Q_summary_1 %>%
  group_by(trial_block) %>%
  summarise_at(vars(Q_pred), list(Q_mean = mean), na.rm=T)

fig_Q_per_trial_1 <- ggplot(Q_summary_per_trial_1, aes(x=trial_block, y=Q_mean)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=12)

fig_Q_per_trial_1

# correct card
Q_summary_2 <- Q_summary[Q_summary$row_odd==0,]

Q_summary_per_trial_2 <- Q_summary_2 %>%
  group_by(trial_block) %>%
  summarise_at(vars(Q_pred), list(Q_mean = mean), na.rm=T)

fig_Q_per_trial_2 <- ggplot(Q_summary_per_trial_2, aes(x=trial_block, y=Q_mean)) +
  geom_line() +
  geom_point() +
  theme_bw(base_size=12)

fig_Q_per_trial_2
```
